Modeling the Linux page cache for accurate simulation of data-intensive applications

### 2020-12-15 paper_io_sim.tex
Fixed conclusion and references

    \section{Conclusion}
    \label{discussion}
    We designed a model of the Linux page cache and implemented it in
    the \simgrid-based \wrench library to simulate distributed
    applications. Evaluation results show that our model very
    substantially improves simulation accuracy, decreasing absolute
    relative simulation errors up to 9$\times$ (see results of
    the single-threaded experiment).
    Our page cache model is publicly available
    in the \wrench GitHub repository.

    Page cache simulation would be instrumental in a number of studies.
    For instance, it is now common for HPC clusters to run 
    application jobs in Linux control groups (cgroups), where resource consumption 
    including memory is controlled.

    With the page cache, I/O times are strongly dependent on the amount of
    available memory, which is itself impacted by the use of anonymous
    memory by applications. A more accurate simulation of application
    memory consumption would therefore be a relevant future direction to
    further improve I/O simulation.

    I/Os remain much more complex than is currently available in simulation
    models.

        Simulation results could be made even more accurate by a deeper
        investigation of page cache flushing and eviction order. 
        The availability of asymmetrical bandwidths in the
        forthcoming \simgrid release will also directly improve our
        simulation accuracy.

        Relevant future work includes simulating
        application memory consumption, random I/O, file readahead and
        persistent disk storage.

        \tristan{discussion could be fleshed out. cgroups}

>>>>

    \section{Conclusion}
    \label{discussion}
    We designed a model of the Linux page cache and implemented it in the
    \simgrid-based \wrench library to simulate distributed applications.
    Evaluation results show that our model improves simulation accuracy
    substantially, decreasing absolute relative simulation errors up to
    9$\times$ (see results of the single-threaded experiment). The
    availability of asymmetrical disk bandwidths in the forthcoming
    \simgrid release will further improve the accuracy of our simulations.
    Our page cache model is publicly available in the \wrench GitHub
    repository.

    Page cache simulation would be instrumental in a number of studies. For
    instance, it is now common for HPC clusters to run applications in
    Linux control groups (cgroups), where resource consumption is limited, including
    memory and therefore page cache usage. Using our simulator,
    it would be possible to study the interaction between memory allocation
    and I/O performance, for instance to improve scheduling algorithms.
    Our simulator could also be utilized to evaluate solutions that
    reduce the impact of network file transfers on distributed
    applications, such as burst
    buffers~\cite{ferreiradasilva-fgcs-bb-2019}, hierarchical file
    systems~\cite{islam2015triple}, active storage~\cite{5496981}, or specific
    hardware architectures~\cite{hayot2020performance}.

    However, I/Os remain much more complex than is currently available in
    simulation models, which could drastically limit the accuracy of simulations.
    Relevant extensions to this work could include more
    accurate descriptions of anonymous memory usage by applications. 
    With the page cache, I/O times are indeed strongly
    dependent on the amount of available memory, which is itself impacted
    by the use of anonymous memory. File access patterns, 
    including random I/O, might also be worth including in application models,
    as they directly impact page cache content.
    % readahead and persistent storage could also be adeed
    % Simulation results could be made even more accurate by a deeper
    % investigation of page cache flushing and eviction order.
